---
title: "ISTA 320 Data Visualization Challenge 05"
author: "Aaron Elledge"
date: "Spring 2021"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
```

For this data viz challenge, you will be working with the same dataset as the [Line Plots Case Study -- 2016 Advanced Placement Classes Part 1](), which was downloaded from [2016 Advanced Placement Exam Scores](https://www.kaggle.com/collegeboard/ap-scores?select=exams.csv)


# Data Wrangling Part 1

In the next code block, make sure you:

1. load the `tidyverse` library
1. read "data/exams.csv" in using `read_csv()`
1. inspect data to get an idea of what it looks like

```{r}

library(tidyverse)
exams <- read_csv("data/exams.csv")
glimpse(exams)
view(exams)
```

# Data Visualization 1

> Question 1: How many students, from `All Students (2016)` (i.e., from all the students that took AP exams in 2016), got each score (from 1 to 5) across different exam subjects.

1. start with the object you created when you `read_csv()`
1. filter out Scores that are "All" and "Average" (keep scores that are from 1 to 5)
1. start a ggplot with the following in mind: score is your sequential variable, `All Students (2016)` is your numeric variable (with student counts), and `Exam Subject` is your categorical variable.
1. use `geom_line` to draw a line plot
1. make any other adjustments to make your plot clearer (e.g., use `facet_wrap`, change scales, add a caption)

```{r fig.height=20}
exams_score <- exams %>%
  filter(Score < 6)
#view(exams_score)

exams_score <- exams_score %>%
  select(`Exam Subject`, Score , `All Students (2016)`)
#view(exams_score)

exams_score %>%
  ggplot(aes(x = factor(Score),
             y = `All Students (2016)`),
         group = `Exam Subject`) +
  geom_line() +
  geom_point() +
  facet_wrap(~ `Exam Subject`, ncol = 3,
             scales = "free")
labs(caption = "How Many Students from 2016 got scores 1-5 for each Subject") +
  scale_y_continuous(expand = expansion(mult = c(.5,.5))) +
  theme(strip.text = element_text(size = 6))

```

# Data Wrangling Part 2

> Question 2: What is the distribution of Average scores across exam subjects and gender (male vs. female)?

First, create a new dataframe with exam subject and average scores for male and female students. 

First you need to filter your original data to keep Score that is equal to "Average" only -- when you do, you are keeping rows that diplay average scores only (instead of student counts).

Then use `select()` to keep the following columns:

- `Exam Subject`
- `Students (Male)`
- `Students (Female)`

```{r}
# create a new data frame that holds the results of:
# start with the original data you read through read_csv and then
# filter the data to keep only Score that is equal to "Average" and then
# select Exam Subject, Students (Male), and Students (Female) 
# (remember to use back tick for column names inside select())
exams_score <- exams %>%
  filter(Score == "Average")
#view(exams_score)

gender <- exams_score %>%
  select(`Exam Subject`, `Students (Male)`, `Students (Female)`)
view(gender)


```

Create a pivoted dataframe, starting with the selected dataframe you just created, and `pivot_longer()` the columns "Students (Male)" and "Students (Female)". Make any changes to the new gender column (e.g., clean it up so it only says "male" or "female").

```{r}
gender_avg <- gender %>%
  pivot_longer(cols = c(`Students (Male)`,
                        `Students (Female)`),
               names_to = c("gender"))

gender_avg <- gender_avg %>%
  mutate(gender = gsub("Students \\(|\\)", "", gender))

gender_avg %>%
  head()


```
You should now have a tidy dataframe with the following three variables: exam subject, gender, and average score.

# Data Visualization 2

To answer question 2, plot a line plot of "Average" scores by student count, across gender:

1. start with the tidy dataframe you created in the previous section
1. start a ggplot with the following in mind: gender is your "sequential" variable, your average score is your numeric variable, and `Exam Subject` is your categorical variable for group mapping.
1. use `geom_line` to draw a line plot
1. use `facet_wrap` to split your plot in subplots by `Exam Subject`
1. make any other adjustments to make your plot clearer (e.g., add a caption, change number of columns in your facet_wrap, add geom_label)


```{r fig.height=30}

gender_avg %>%
  ggplot(aes(x = factor(gender), y = value, group = `Exam Subject`)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ `Exam Subject`, ncol = 3, scales = "free") +
  labs(caption = "Average scores by student count across gender") +
  scale_y_continuous(expand = expansion(mult = c(.5,.5)))
```

