---
title: "ISTA 320 Data Visualization Challenge 06"
author: "Aaron Elledge"
date: "Spring 2021"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
library(tidyverse)

```

For this data viz challenge, you will be working with the GME (GameStop) stock price dataset from [Yahoo Finance](https://finance.yahoo.com/quote/GME/history?period1=1583193600&period2=1614729600&interval=1d&filter=history&frequency=1d&includeAdjustedClose=true)

The data is inside the `data` folder, and it's called `GME.csv`. Alternatively you can download a more up-to-date version of the data from the link provided above.

```{r}
dir("data")
gme <- read.csv("data/GME.csv")

View(gme)
head(gme)

```

# Read the Data in

```{r}
# Data Analysis

#Plot a timeseries (i.e., you need to map a `date` variable in your plot).


gme %>%
  ggplot(aes (x = Date,
             y = Close))+
  geom_point()+
  geom_line()
```


```

ENTER YOUR INTEPRETATION OF YOUR PLOT HERE

## My interpretation is the closing prices over time, as it becomes later on the closing prices rise and then drops. Overall, the closing price doesn't change much until later on.